The Second Data Release of the INT Photometric H$\alpha$ Survey of the Northern Galactic Plane (IPHAS DR2)

### 2014-03-14 paper/iphas-dr2.tex
Typos

\author[G. Barentsen
et. al]{Geert Barentsen$^{1}$\thanks{E-mail:geert@barentsen.be},
H. J. Farnhill$^1$,
J. E. Drew$^1$,
E. A. Gonz$\acute{\rm{a}}$lez-Solares$^2$, \newauthor
R. Greimel$^3$,
M. J. Irwin$^2$,
B. Mizalski$^{99}$,
C. Ruhland$^1$,
P. Groot$^4$,
A. Mampaso$^5$, \newauthor
S. E. Sale$^6$,
M. J. Barlow$^{99}$,
R. L. M Corradi$^{99}$,
J. J. Drake$^{99}$,
J. Fabregat$^{99}$, \newauthor
B. T. Gaensicke$^{99}$,
A. S. Hales$^{99}$,
J. Irwin$^{99}$,
C. Knigge$^{99}$,
D. J. Lennon$^{99}$, \newauthor
J. R. Lewis$^{99}$,
M. Mohr-Smith$^{99}$,
R. A. H. Morris$^{99}$,
T. Naylor$^{99}$,
Q. A. Parker$^{99}$, \newauthor
S. Phillipps$^{99}$,
R. Raddi$^{99}$,
P. Rodriguez-Gil$^{99}$,
L. Sabin$^{99}$,
D. Steeghs$^{99}$, \newauthor
Y. C. Unruh$^{99}$,
K. Viironen$^{99}$,
J. S. Vink$^{99}$,
N. A. Walton$^{99}$,
N. J. Wright$^{99}$, \newauthor
A. A. Zijlstra$^{99}$,
plus those I forgot (sorry).
\newauthor\\
$^{1}$School of Physics, Astronomy \& Mathematics, University of Hertfordshire, College Lane, Hatfield, Hertfordshire, AL10 9AB, U.K.\\
$^{2}$Institute of Astronomy, Cambridge University, Madingley Road, Cambridge, CB3 OHA, U.K.\\
$^{3}$IGAM, Institute of Physics, University of Graz, Universit\"atsplatz 5, Graz, Austria.\\
$^{4}$Afdeling Sterrenkunde, Radboud Universiteit Nijmegen, Faculteit NWI, Postbus 9010, 6500 GL Nijmegen, The Netherlands.\\
$^{5}$Instituto de Astrof\'isica de Canarias, 38200 La Laguna, Tenerife, Spain.\\
$^{6}$Rudolf Peierls Centre for Theoretical Physics, Keble Road, Oxford, OX1 3NP, U.K.\\
$^{99}$Please add your current affiliation.\\
}

\begin{document}
\date{Current draft typeset \today}
\pagerange{\pageref{firstpage}--\pageref{lastpage}} \pubyear{2013}

@@

\begin{abstract} % BACKGROUND, OBJECTIVE, METHODS, RESULTS, CONCLUSIONS
The INT/WFC Photometric H$\alpha$ Survey 
of the Northern Galactic Plane (IPHAS)
is a 1800 deg$^2$ imaging survey
covering the entire northern Milky Way at $|b| < 5^\circ$
in the $r$, $i$ and \ha\ filters 
using the Wide Field Camera (WFC) 
on the 2.5-meter Isaac Newton Telescope (INT) in La Palma.
We present the first quality-controlled and
uniformly-calibrated source catalogue
to have been extracted from the images,
providing single-epoch photometry
for 219 million unique sources
across 92\% of the survey area.
The observations were carried out between 2003 and 2013
at a median seeing of 1.1 arcsec
and to a depth of $r=21.2\pm 0.5$, $i=20.0\pm 0.3$ and \ha$=20.3\pm 0.3$
($5\sigma$ limits, Vega system).
We explain the data reduction 
and quality control procedures,
describe and test the new uniform calibration,
and detail the construction of the 97-column source catalogue.
We find that the calibration
and the photometric repeatability are accurate to
0.03~mag (rms),
and provide a series of recommended quality criteria
which can be used to select the most reliable data
from the catalogue.
Finally, we demonstrate the ability of the 
catalogue's unique
$(r-i)$/($r-\ha)$ colour-colour diagram to
(i) characterise stellar populations and extinction regimes
towards different Galactic sight-lines
and (ii) select candidate \ha\ emission-line objects.
IPHAS is the first survey to offer comprehensive CCD photometry
of point sources across the Galactic Plane at visible wavelengths,
providing the much-needed counterpart
to several existing near- and mid-infrared surveys,
and providing images for some of the most
crowded regions to be faced by Gaia.
\end{abstract}

@@

The catalogue is made available through the Vizier
catalogue tool (http://vizier.u-strasbg.fr),
where it is known as the ``IPHAS DR2 Source Catalogue''
and where it can be queried
(catalogue ID ??? Conesearch/TAP URL??).
The catalogue can also be downloaded in its entirety
from the IPHAS website (www.iphas.org) as a collection 
of binary FITS tables which comprise 50\,GB,
together with a script
to ingest these tables into a PostgreSQL database.
Finally, the website also provides access to the pipeline-processed
imaging data, which we have updated to include
the re-calibrated DR2 zeropoint in the image headers.

@@

The observations included in this release
achieved a median seeing of 1.1 arcsec
and a $5\sigma-$depth of $r=21.2\pm 0.5$, $i=20.0\pm 0.3$, \ha$=20.3\pm 0.3$.
The global calibration and photometric repeatability
were found to be accurate at the level of $\sigma=0.03$ mag,
providing a significant improvement over the 
previous data release.
The source catalogue provides the best-available
single-epoch astrometry and photometry
for 219~million unique sources.
For 86 million of these objects the catalogue
provides photometry which can be considered to be reliable
in all three bands.

In future work, we aim to draw upon the PanSTARRS photometric
survey to further improve the accuracy of our global calibration.
We will also aim to correct the IPHAS photometry for the
radial field distortions,
and aim to produce a catalogue
which details all the secondary detections
to aid time-domain studies.

The data-taking strategy developed for IPHAS
has since been reapplied to carry out a companion INT/WFC survey called UVEX
in U/$g$/$r$/He {\sc i} \citep{Groot2009}
and a southern companion survey in $u$/$g$/$r$/$i$/\ha\ 
called VPHAS+ \citep{Drew2014}.
We intend to use the present work
as a template for future releases
to be prepared for these companion surveys,
allowing new insights into the contents and structure 
of our own backyard, the Milky Way, to be revealed.

\section*{Acknowledgements}

>>>>

\author[G. Barentsen
et. al]{G. Barentsen$^{1}$\thanks{E-mail:geert@barentsen.be},
H. J. Farnhill$^1$,
J. E. Drew$^1$,
E. A. Gonz$\acute{\rm{a}}$lez-Solares$^2$, 
R. Greimel$^3$, \newauthor
M. J. Irwin$^2$,
B. Mizalski$^{4}$,
C. Ruhland$^1$,
P. Groot$^5$,
A. Mampaso$^6$, 
S. E. Sale$^7$, \newauthor
M. J. Barlow$^{8}$,
R. L. M Corradi$^{6}$,
J. J. Drake$^{9}$,
J. Fabregat$^{10}$, 
B. T. Gaensicke$^{11}$, \newauthor
A. S. Hales$^{12}$,
J. Irwin$^{9}$,
C. Knigge$^{13}$,
D. J. Lennon$^{14}$, 
J. R. Lewis$^{2}$,
M. Mohr-Smith$^{1}$, \newauthor
R. A. H. Morris$^{15}$,
T. Naylor$^{16}$,
Q. A. Parker$^{17}$, 
S. Phillipps$^{13}$,
R. Raddi$^{11}$, \newauthor
P. Rodriguez-Gil$^{6}$,
L. Sabin$^{18}$,
D. Steeghs$^{11}$, 
Y. C. Unruh$^{19}$,
K. Viironen$^{20}$, \newauthor
J. S. Vink$^{21}$, 
N. A. Walton$^{2}$,
N. J. Wright$^{1}$,
A. A. Zijlstra$^{22}$.
\newauthor\\
$^{1}$School of Physics, Astronomy \& Mathematics, University of Hertfordshire, College Lane, Hatfield, Hertfordshire, AL10 9AB, U.K.\\
$^{2}$Institute of Astronomy, Cambridge University, Madingley Road, Cambridge, CB3 OHA, U.K.\\
$^{3}$IGAM, Institute of Physics, University of Graz, Universit\"atsplatz 5, Graz, Austria.\\
$^{4}$South African Astronomical Observatory, P.O. Box 9, Observatory, 7935 Cape Town, South Africa.\\
$^{5}$Afdeling Sterrenkunde, Radboud Universiteit Nijmegen, Faculteit NWI, Postbus 9010, 6500 GL Nijmegen, The Netherlands.\\
$^{6}$Instituto de Astrof\'isica de Canarias, 38200 La Laguna, Tenerife, Spain.\\
$^{7}$Rudolf Peierls Centre for Theoretical Physics, Keble Road, Oxford, OX1 3NP, U.K.\\
$^{8}$University College London, Department of Physics \& Astronomy, 
Gower Street, London WC1E 6BT, U.K.\\
$^{9}$Harvard-Smithsonian Center for Astrophysics, 60 Garden Street, 
Cambridge, MA 02138, U.S.A. \\
$^{10}$Observatorio Astr\'onomico, Universidad de Valencia,
Catedr\'atico Jos\'e Beltr\'an 2, 46980 Paterna, Spain\\
$^{11}$Department of Physics, University of Warwick, Gibbet Hill Road, Coventry, CV4 7AL, U.K.\\
$^{12}$ Joint ALMA Observatory, Alonso de Córdova 3107, Vitacura 763-0355, Santiago, Chile. \\
$^{13}$School of Physics \& Astronomy, University of Southampton, Southampton, SO17 1BJ, U.K.\\
$^{14}$European Space Astronomy Centre (ESAC), Villafranca del Castillo, Villanueva de la Canada, E-28692 Madrid, Spain.\\
$^{15}$School of Physics, Bristol University, Tyndall Avenue, Bristol, BS8 1TL, U.K.\\
$^{16}$School of Physics, University of Exeter, Stocker Road, Exeter, EX4 4QL, U.K.\\
$^{17}$Department of Physics \& Astronomy, Macquarie University, NSW 2109, Australia\\
$^{18}$Instituto de Astonom\'ia y Meteorolog\'ia, Departamento de F\'isica, CUCEI, Universidad de Guadalajara, Mexico.\\
$^{19}$Department of Physics, Blackett Laboratory, Imperial College
London, Prince Consort Road, London, SW7 2AZ, U.K.\\
$^{20}$Centro de Estudios de F\'isica del Cosmos de Arag\'on, Plaza San Juan 1, Planta 2, Teruel, 44001, Spain. \\
$^{21}$Armagh Observatory, College Hill, Armagh, Northern Ireland,
BT61 9DG, U.K.\\
$^{22}$Jodrell Bank Centre for Astrophysics, School of Physics \&
Astronomy, University of Manchester, Oxford Road, Manchester M13 9PL, U.K.}

\begin{document}
\date{Current draft typeset \today}
\pagerange{\pageref{firstpage}--\pageref{lastpage}} \pubyear{2014}

@@

\begin{abstract} % BACKGROUND, OBJECTIVE, METHODS, RESULTS, CONCLUSIONS
The INT/WFC Photometric H$\alpha$ Survey 
of the Northern Galactic Plane (IPHAS)
is a 1800 deg$^2$ imaging survey
covering the entire northern Milky Way at $|b| < 5^\circ$
in the $r$, $i$ and \ha\ filters 
using the Wide Field Camera (WFC) 
on the 2.5-meter Isaac Newton Telescope (INT) in La Palma.
We present the first quality-controlled and
uniformly-calibrated source catalogue
to have been extracted from the images,
providing single-epoch photometry
for 219 million unique sources
across 92\% of the survey area.
The observations were carried out between 2003 and 2013
at a median seeing of 1.1 arcsec
and to a depth of $r=21.2\pm 0.5$, $i=20.0\pm 0.3$ and \ha$=20.3\pm 0.3$
($5\sigma$ limits, Vega system).
We explain the data reduction 
and quality control procedures,
describe and test the new uniform calibration,
and detail the construction of the 97-column source catalogue.
We show that the calibration
and the photometric repeatability are accurate to
0.03~mag (rms),
and provide a series of recommended quality criteria
which can be used to select the most reliable data
from the catalogue.
Finally, we demonstrate the ability of the 
catalogue's unique
$(r-i)$/($r-\ha)$ colour-colour diagram to
(i) characterise stellar populations and extinction regimes
towards different Galactic sight-lines
and (ii) select candidate \ha\ emission-line objects.
IPHAS is the first survey to offer comprehensive CCD photometry
of point sources across the Galactic Plane at visible wavelengths,
providing the much-needed counterpart
to infrared surveys,
whilst also providing images for some of the most
crowded regions to be faced by Gaia.
\end{abstract}

@@

The catalogue is made available through the Vizier
catalogue tool (http://vizier.u-strasbg.fr),
where it is known as the ``IPHAS DR2 Source Catalogue''
and where it can be queried
(catalogue ID ??? Conesearch/TAP URL??).
The catalogue can also be downloaded in its entirety
from the IPHAS website (www.iphas.org) as a collection 
of binary FITS tables which comprise 50\,GB,
together with a script
to ingest these tables into a PostgreSQL database.
Finally, the website also provides access to the pipeline-processed
imaging data, which we have updated to include
the re-calibrated zeropoints and the best-available
astrometric solution in the image headers.

@@

The observations included in this release
achieved a median seeing of 1.1 arcsec
and a $5\sigma-$depth of $r=21.2\pm 0.5$, $i=20.0\pm 0.3$, \ha$=20.3\pm 0.3$.
The global calibration and photometric repeatability
were found to be accurate at the level of $0.03$ mag (rms),
providing a significant improvement over the 
previous data release.
The source catalogue provides the best-available
single-epoch astrometry and photometry
for 219~million unique sources.
We detailed a list of recommended quality criteria
to select objects with reliable colours from the catalogue.
Finally, we demonstrated the use of the survey's unique
$(r-i)$/$(r-\ha)$ diagram for characterising stellar populations
and selecting emission-line objects.

In future work, we intend to draw upon the PanSTARRS photometric
survey to further improve the accuracy of our global calibration.
We will also aim to correct the IPHAS photometry for the
radial field distortions,
and plan to produce a catalogue 
which details all the secondary detections
to aid time-domain studies.

The data-taking strategy developed for IPHAS
has since been reapplied to carry out a companion INT/WFC survey called UVEX
in U/$g$/$r$/He {\sc i} \citep{Groot2009}
and a southern survey in $u$/$g$/$r$/$i$/\ha\ 
called VPHAS+ \citep{Drew2014}.
We intend to use the present work
as a template for generating catalogues
from these companion surveys.
When combined, these surveys
will offer new insights
into the contents and structure of the Milky Way
across 360$^\circ$.

\section*{Acknowledgments}
